Entrainment of low Mach number thermals in stratified domains

### 2019-04-30 tex/stratified_thermals.tex
gets paper writing actually started

\received{October 19, 2018}
\revised{\today}
\accepted{??}%\today}
\submitjournal{ApJ}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% TITLE & AUTHORS
\shorttitle{Entrainment of thermals in stratified atmospheres}
\shortauthors{Anders et al.}

\begin{document}
\title{Entrainment of low Mach number thermals in stratified atmospheres}

@@

%%%%% Body of the paper
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% INTRODUCTION
\section{Introduction}
\label{sec:intro}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% EXPERIMENT SECTION
\section{Experiment} 
\label{sec:experiment}

We perform direct numerical simulations of the evolution of dry thermals
by solving the fully compressible Navier-Stokes equations,
\begin{gather}
\frac{D \ln\rho}{Dt} = -\Div{\bm{u}}
	\label{eqn:continuity}
\\
\frac{D \bm{u}}{D t} =
-\frac{1}{\rho}\grad P + \bm{g} + \Div{\stressT}, 
	\label{eqn:dim_bouss_momentum}
\\
\frac{D T}{D t} + (\gamma - 1)T\Div{\bm{u}} + \frac{1}{\rho c_V}\Div{-\rho\chi\grad T} = 
\frac{1}{\rho c_V}(\stressT \cdot \grad)\cdot \bm{u},
	\label{eqn:dim_bouss_energy}
\end{gather}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% RESULTS SECTION
\section{Results}
\label{sec:results}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% CONCLUSION SECTION
\section{Discussion}
\label{sec:discussion}

@@

\appendix
\section{Table of Simulations}
\label{appendix:table}

>>>>

\received{\today}
\revised{\today}
\accepted{??}%\today}
\submitjournal{ApJ}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% TITLE & AUTHORS
\shorttitle{Stratified Thermals}
\shortauthors{Anders et al.}

\begin{document}
\title{Entrainment of low Mach number thermals in stratified domains}

@@

%%%%% Body of the paper
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% INTRODUCTION
\section{Introduction}
\label{sec:intro}
(Para 1) Brief intro to solar convective conundrum.

(Para 2) Entropy rain history and expansion by Axel

(Para 3) Boussinesq thermals and their filling factor vs. depth.

(Para 4) What we're doing here, how it fits into this context.

(Para 5) Standard last paragraph of an intro?

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% THEORY SECTION
\section{Theory}
\label{sec:theory}
The evolution of thermals as buoyant vortex rings has been well described in
the unstratified, Boussinesq limit for decades (as early as e.g., CITE, and
see \citet{lecoanet&jeevanjee2018} for other sources). Buoyant motions 
in the atmospheres of stars and planets are generally large enough to feel the
atmospheric stratification, and therefore a more thorough tratment of the
evolution of thermals in stratified domains is required to understand the nature
of thermal entrainment in nature.

In this work, we focus on the non-dissipative, low Mach number regime, 
in which the ideal anelastic equations are a decent approximation to the fully
compressible equations. In this regime, the buoyancy is directly proportional to the
specific entropy. In the absence of diffusion, or in the limit where diffusivities
are sufficiently small, the volume-integrated total entropy is constant,
\begin{equation}
B \equiv \int_{\mathcal{V}} \rho\, s\, dV = \text{const},
\end{equation}
where $s = c_V \ln T - R \ln\rho$ is the specific entropy, with $c_V$ the
specific heat at constant volume and $R$ the ideal gas constant.

In a stratified domain, the hydrodynamic impulse is defined
\citet{shivamoggi2010},
\begin{equation}
\bm{I} = \frac{1}{2}\int_{\mathcal{V}} \bm{x}\times(\grad\times(\rho\bm{u}))dV,
\end{equation}
where $\bm{x}$ is the position vector, $\bm{u}$ is vector velocity, $\rho$ is density,
and $\mathcal{V}$ is the volume being integrated over. Furthermore, changes in the impulse
can be expressed
\begin{equation}
\frac{\partial\bm{I}}{\partial t} = \int_{\mathcal{V}}\frac{\partial}{\partial t}(\rho \bm{u})dV
= B\hat{z} + S,
\end{equation}
where $S$ is a combination of surface terms that disappears upon appropriate boundary conditions
on $\mathcal{V}$ \citep{shivamoggi2010}. As we have assumed that the buoyancy, $B$, is constant,
we can straightforwardly integrate
\begin{equation}
I_z = B t + I_0,
\end{equation}
for some constant $I_0$.

(need to work through this section carefully).
Assuming an adiabatically stratified atmosphere in
hydrostatic equilibrium, the vertical momentum equation is
\begin{equation}
 = - \partial_z \varpi +  g\frac{S_1}{c_P},
\label{eqn:buoyancy}
\end{equation}
where $w$ is vertical velocity, $S_1$ is specific entropy, $\varpi$ is the reduced pressure,
$D/Dt = \partial_t + (\bm{u}\cdot\grad)$ is the (eulerian? lagrangian?)
derivative, $g$ is gravity, and $c_P$ is the specific heat at constant pressure.
In the anelastic approximation, the density stratification is constant in time. 
OH GOD I DON'T KNOW THIS MATH WELL MAGIC MAGIC MAGIC.
By doing magic, we arrive at
\begin{equation}
P_z = \beta B t + P_0.
\end{equation}

Under the assumption that the thermal develops into a thin-core 
propagating vortex ring whose vortex core is radius $r$ away from the axis of symmetry, 
the impulse can be approximated as $I_z \approx \pi \rho r^2 \Gamma$, where 
$\Gamma$ is the circulation of the thermal vortex, which we assume to be constant. 
Rearranging, we find our first result,
\begin{equation}
r = \sqrt{\frac{B t + I_0}{\pi\rho\Gamma}}.
\label{eqn:r_theory}
\end{equation}
In the limit where $\rho \rightarrow \text{constant}$, as in the Boussinesq regime,
we retrieve the $r \propto \sqrt{t}$ scaling found in the Boussinesq regime by
\citet{lecoanet&jeevanjee2018}. We find that the inclusion of stratification adds the
additional complexity of $r \propto \rho^{-1/2}$, such that downward-propogating 
vortex rings (as studied here) will not entrain to the same degree as boussinesq thermals,
and upward-propogating rings will entrain more.

We further assume that the volume-integrated momentum, $P_z \approx \rho V w_{th}$,
where $w_{th}$ is the vertical velocity of the thermal as a whole, and the volume
of the fluid region propogating with the thermal is a spheroid, $V = V_0 r^3$, for
some constant $V_0$. Plugging Eqn.~\ref{eqn:r_theory} into this approximation, we
find that
\begin{equation}
\rho^{-1/2}w = \left(\frac{(\pi\Gamma)^{3/2}}{V_0}\right)\frac{\beta B t + M_0}{(B t + I_0)^{3/2}}.
\end{equation}
Substituting $w = dz/dt$ and $\tau \equiv (Bt + I_0)/\Gamma$, We retrieve
\begin{equation}
\frac{dz}{\sqrt{\rho(z)}} 
= \left(\frac{\pi^{3/2}\Gamma}{V_0}\right)
\left(\beta\Gamma\tau^{-1/2} + [M_0 - \beta I_0]\tau^{-3/2}\right)d\tau.
\end{equation}
If the atmospheric stratification in which the thermal is falling is known, this
result can be straightforwardly integrated with $\rho(z)$ plugged in in order to
find the position of the thermal versus time. While we will leave this result
general for now, we will plug in our polytropic stratification at the end of
section \ref{sec:experiment}, and show that the resulting expression does a
good job of explaining the evolution of thermals in these atmospheres in
section \ref{sec:results}.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% EXPERIMENT SECTION
\section{Experiment} 
\label{sec:experiment}
blah blah blah polytrope blah blah blah (do it with a general length scale).

blah blah blah 2D anelastic blah blah blah, nondimm'd for a direct comparison with
boussinesq.

blah blah blah 3D Fully compressible blah blah blah

blah blah blah thermal initial conditions, pressure equilibrium blah blah blah.

blah blah blah comparison of 2D and 3D blah blah blah. z v t, r v z, w v z. Also side-by-side
pictoral comparison with diff.

blah blah blah T(t) and derivative blah blah blah.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% RESULTS SECTION
\section{Results}
\label{sec:results}
Picture of a big old grid of z v t, r v z, w v z.

Pretty picture showing thermal evolution (comparing low and high stratification).

Verification of theory by simulations

Table of found parameters for the fits

Two regimes: stalling and falling.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% CONCLUSION SECTION
\section{Discussion}
\label{sec:discussion}
Wild speculation about extensions to the solar regime. Do things on the sun shrink
to the point where they viscously dissipate?

Talk about what would happen if we were to study up-thermals.

Extensions, and the fact that we trust these results will likely hold in the solar regime.

@@

\appendix
\section{Thermal Tracking}
\label{appendix:tracking}

\section{Table of Simulations}
\label{appendix:table}
